---
title: "autoshiny -- R package for automatic transformation of an R function into a Shiny app"
author: "Aleksander Rutkowski"
date: "`r format(Sys.Date())`"
output: github_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Rendering this doc:
# rmarkdown::render(paste0(getwd(),'/README.Rmd'), envir=globalenv())
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(engine='R')
options(python_path = "C:/Python27/python.exe")
i <- 0
print.shiny.appobj2 <- function(x, ...) { # needed to prepare the screenshots, overriding shiny's
    i <<- i+1
    fname <- paste0('screenshot',i,'.png')
    cat("", file=fname) # pre-save the file placeholder so that pandoc dones not complain
    webshot::appshot(structure(x, class='shiny.appobj'),
                     file=paste0(getwd(),'/',fname),
                     vheight=300,
                     delay=3)
    cat(paste0('\n![](https://cdn.rawgit.com/alekrutkowski/autoshiny/master/',fname,')\n\n', # needs results='asis' in chunk options
               'See the [source code](https://github.com/alekrutkowski/autoshiny/tree/master/Example_',i,') generated with `makeFiles` ',
               '(and formatted with [rfmt](https://github.com/google/rfmt)) ',
               'of the app whose initial-state screenshot is displayed above.')) 
}
```

## Installation

[From CRAN](https://CRAN.R-project.org/package=autoshiny):

```{r, eval=FALSE}
install.packages('autoshiny')
```

or the latest version from GitHub:

```{r, eval=FALSE}
# if package `devtools` not installed, first do this:
# install.packages('devtools')
devtools::install_github('alekrutkowski/autoshiny')
```

## Key info

There are two key twin functions: `makeApp` and `makeFiles`.
Both of them take a function as their first argument/parameter.
Function `makeApp` returns [a Shiny app object](https://rdrr.io/cran/shiny/man/shinyApp.html).
`makeFiles` produces `ui.R` and `server.R` files.
These files can be further edited to tweak the app if needed.

Using **autoshiny** does not imply any run-time dependency of the compiled app on **autoshiny**
i.e. **autoshiny** is needed only at the compile time. **autoshiny** uses standard Shiny
input and output widgets and render functions. Tiny helper functions are embedded in the
compiled app code to make it self-contained.

All the arguments/parameters of the function passed to `makeApp` and `makeFiles` *must have default values* which will
be used by **autoshiny** to define each argument's:

- type/class -- Shiny input widget type,
- allowed values -- Shiny input widget's allowed states,
- pre-selected/start-up value of the Shiny input widget -- for categorical values (e.g. integers, factors, character vectors
that will be the first element of the vector).

The default values of the function arguments/parameters will be also used to pre-evaluate that function in order to
test it and to determine the type of its output (return value / side effect) and, hence, the Shiny output widget.

## Examples

```{r, eval=FALSE}
library(autoshiny)
```

```{r}
library(shiny)
```


```{r, include=FALSE}
j <- 0
makeApp <- function(...) {
    j <<- j + 1
    autoshiny::makeFiles(...,directory=getwd())
    ExN <- paste0('Example_',j)
    if (!dir.exists(ExN))
        dir.create(ExN)
    fls <- c('server.R','ui.R')
    fls2 <- paste0(ExN,'/',fls)
    file.copy(fls, fls2)
    file.remove(fls)
    lapply(fls2,
           rfmt::rfmt)
    structure(autoshiny::makeApp(...),
              class='shiny.appobj2')
}
File <- autoshiny::File
```

### Example 1: Trivial anonymous function

```{r, results='asis'}
makeApp(function(x=1:3, y=5:9) x+y)
```

### Example 2: Nicer function and argument names

```{r, results='asis', fig.keep='none'}
`Histogram for normal distribution` <-
    function(`Number of observations` =
                 as.integer(c(100,10,1000))) # as.integer => the argument interpreted as categorical
        plot(hist(rnorm(`Number of observations`))) # Generic R plots as "return values" are supported

makeApp(`Histogram for normal distribution`)
```

### Example 3: Data frame in (upload CSV), data frame out (displayed and downloadable as CSV)

```{r, results='asis'}
`Table of sin and cos values` <-
    function(`Upload CSV file with column "x"` =
                 data.frame(x = seq(0, 2*pi, .25))) {
        dta <- `Upload CSV file with column "x"`
        data.frame(X = dta$x,
                   `Sin of X` = sin(dta$x),
                   `Cos of X` = cos(dta$x),
                   check.names = FALSE)
    }
makeApp(`Table of sin and cos values`)
```

### Example 4: Arbitrary input and output files

```{r, results='asis'}
openxlsx::write.xlsx(data.frame(x=1:5,
                                y=11:15),
                     'my_test_file.xlsx')
`Excel file in and out` <-
    function(`Input Excel file` =
                 File('my_test_file.xlsx')) { # File() obligatory here!
        my.data <- openxlsx::read.xlsx(`Input Excel file`)
        my.data2 <- within(my.data,
                           z <- x + y)
        openxlsx::write.xlsx(my.data2,
                             'my_test_file_2.xlsx')
        File('my_test_file_2.xlsx') # File() obligatory here too!
    }
makeApp(`Excel file in and out`)
```

### Example 5: Using a button as a (re-)evaluation trigger

Use this option if:

- the evaluation of your functon takes time, so it should not be re-evaluated with every minor change of the value of inputs/arguments/parameter;
- the function is impure e.g. depends on some external data fetched internally and takes no arguments/parameters -- in such a case the function would be re-evaluated only through page refresh of the browser; the button is a faster and a more elegant solution.

```{r, results='asis'}
`Get "GDP and main components" from Eurostat` <-
    function() {
        # Getting data from
        # http://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&file=data%2Fnama_10_gdp.tsv.gz
        x <- eurodata::importData('nama_10_gdp')
        head(x, 10)
    }
makeApp(`Get "GDP and main components" from Eurostat`,
        withGoButton = TRUE)
```

### Example 6: Lists of inputs (arguments) and the output list (composite return value) are always decomposed

```{r, results='asis'}
`A function with lists everywhere` <-
    function(`First argument group,` = list(`number one` = 1:3,
                                           `number two` = letters[1:3]),
             `2nd arg group,` = list(`1st argument` = 11:14,
                                    `second arg.` = LETTERS[1:5]))
        list(`Some text` =
                 as.character(c(`First argument group,`$`number two`,
                              `2nd arg group,`$`second arg.`)),
             `Some numbers` =
                 `First argument group,`$`number one` +
                 `2nd arg group,`$`1st argument`,
             `Even a ggplot2 chart` =
                 ggplot2::qplot(a,b,data=data.frame(a=1:20,b=log(1:20))))
makeApp(`A function with lists everywhere`)
```





